The Hot Hand Fallacy in Call of Duty Search and Destroy

### 2022-08-29 paper/paper.tex
everything but table 7 analysis

In professional CoD SnD, the two teams\footnote{In the 2020 season,
  teams played with five players each; in the 2021 and 2022 seasons,
  teams played with four players each.} take turns playing offense and
defense every round. The first to six round wins (best-of-11 round
format) is declared the series winner.\footnote{A short list of rules
  that govern timing follows.

  \begin{itemize}
  \tightlist
  \item
    Each round has a 90 second time limit, not counting the potential
    extension granted if a bomb is planted
  \item
    A bomb plant takes five seconds. The timer resets if the player
    stops planting site prior to completing it.
  \item
    A bomb defuse takes seven seconds. The timer resets if the player
    ``drops'' the bomb.
  \item
    Once the bomb is planted, the round timer is set to 45 seconds.
  \end{itemize}} A round can end in one of four ways:

@@

The data set consists of all SnD matches played in major tournaments and
qualifiers during the past three years\footnote{CoD has roughly gone
  through three eras of professional gaming: (1) Major League Gaming
  (MLG) tournaments prior to 2016; (2) the CoD World League (CWL),
  initiated in 2016; and (3) the 12-franchise CoD League (CDL),
  operating since 2020. The CDL has completed three year-long
  ``seasons'' as of August 2022.

  CoD is fairly unique compared to other esports in that it runs on an
  annual lifecycle, with releases coming in the late fall. A new game
  under the same brand---Call of Duty---is published every year by a
  rotating set of developers. Each new game bears resemblance to past
  ones, often introducing relatively small variations (``improvements'')
  to graphics, game modes, and other facets of gameplay. During the CDL
  era, the games released have been Modern Warfare (2020), Cold War
  (2021) and Vanguard (2022).}, totaling 7,792 rounds across 852
series.\footnote{288 of the series occur in major tournaments, which are
  considered to be more competitive than the qualifiers since they are
  played in person (COVID-permitting), eliminating randomness due to
  online latency and ping. The qualifiers are played online.} Data has
been collected in spreadsheets by community member
``IOUTurtle''.\footnote{Raw data: \url{https://linktr.ee/CDLArchive}.
  Cleaned data:
  \url{https://github.com/\%7Bauthor\%7D/\%7Brepo\%7D/blob/master/data/\%7Bfile\%7D.csv}.}

Sweeps (6-0 series) make up 4.7\% of all series, while 6-5 series make
up 23.7\% of series. The observed offensive round win percentage across
all rounds is \(\bar{\tau}\) = 47.8\%.\footnote{Offensive round win
  percentage has been nearly constant across the three games during the
  CDL era: 1. 47.2\% in MW (2020) 2. 47.9\% in Cold War (2021) 3. 48.1\%
  in Vanguard (2022)} Table \ref{tbl:o-win-prop-by-series-state} shows
offensive round win percentages by series ``state'', i.e.~the number of
round wins by each team prior to an upcoming round. Offensive round win
percentage is not quite constant, although never veers more than 10\%
from \(\bar{\tau}\).

@@

The professional competition format balances the frequency with which
teams play offense or defense to start a series, to the extent possible.
Thus, win percentages in certain states are subjected to minimal
selection bias, i.e.~bias due to better teams playing defense more
frequently.\footnote{At tournaments, higher-seeded (``better'') teams
  choose whether they want to start on offense in the SnD series, played
  as the second or fifth game mode of the best-of-five matchup. Their
  opponent is assigned to start on offense in the SnD slot not chosen.
  Anecdotally, the better team tends to choose to play defense in the
  first round of the SnD series played as the second game mode in the
  best-of-five matchup, although this choice is not consistent across
  teams, or even with the same team over time.

@@

  This format has minimized the range of rates at which teams start
  series playing offense. The highest rate at which a team has started
  their series playing defense across all tournament and qualifier SnD
  series in a single season is 66\% (31 of 47). The lowest is 38\% (16
  of 42).}

@@

There have been a handful of studies of the distribution of games played
in a series of a professional sport. Most assume a constant probability
\(\phi\) of a given team winning a game in the series, regardless of the
series state. Mosteller (1952) observed that the American League had
dominated the National League in MLB World Series matchups through 1952,
implying that games should not be modeled with a constant
\(\phi_0 = 0.5\). Mosteller proposed three approaches for identifying
the optimal constant probability value of the stronger team in the World
Series, finding \(\hat{\phi} \approx 0.65\) in each case: (1) solve for
\(\phi\) from the observed average number of games won by the loser of
the series, which he called the ``method of moments'' approach; (2)
maximize the likelihood that the sample would have been drawn from a
population in which the probability of a team winning a game is constant
across the series (i.e.~maximum likelihood), and (3) minimize the
chi-squared goodness of fit statistic \(\chi^2\) as a function of
\(\phi\).

Chance (2020) re-examined the constant probability notion in the MLB
World Series (1923--2018), the NBA Finals (1951--2018), and the NHL
Stanley Cup (1939--2018). Chance found strong evidence against the null
hypothesis of \(\phi_0 = 0.5\) in the MLB and NHL championship series
when applying Mosteller's first and second methods.

The concept of momentum goes hand-in-hand with a discussion of the
nature of series outcomes.\footnote{We often use ``streaks'' and
  momentum interchangeably, but as Steeger et al. (2021) note, momentum
  implies dependence between events, whereas streaking does not.} Two
opposing fallacies are observed in the context of momentum: the
``gambler's fallacy'' (negative recency) and the ``hot hand fallacy''
(positive recency). Per Ayton et al. (2004), negative recency is ``the
belief that, for random events, runs of a particular outcome \ldots{}
will be balanced by a tendency for the opposite outcome'', while
positive recency is the expectation of observing future results that
match recent results.

Studying both player streaks and team streaks in basketball, in both
observational and controlled settings, Gilovich et al. (1985),
henceforth GVT, did not find evidence for the hot hand phenomenon.
However, Miller and Sanjurjo (2018), henceforth MS, provided a framework
for quantifying streak selection bias, which effectively works in the
manner posited by the gambler's fallacy. Specifically, MS stated that a
``bias exists in a common measure of the conditional dependence of
present outcomes on streaks of past outcomes in sequential data''
implying that, under i.i.d. conditions, ``the proportion of successes
among the outcomes that immediately follow a streak of consecutive
successes is expected to be strictly less than the underlying
(conditional) probability of success''. When applying streak selection
bias to GVT's data, MS came to the opposite conclusions as GVT.

@@

Despite the plethora of existing research on games played in a series
and momentum in sports, these topics have yet to be investigated
heavily, if at all, in esports. Work has been done to examine
intra-round, second-by-second win probability in other FPS titles such
as Counter-Strike (Xenopoulos, Freeman, and Silva 2022) and Valorant
(DeRover 2021), both of which are round-based like CoD SnD. However,
considering Counter-Strike and Valorant specifically, research on
round-level trends seems non-existent, perhaps for one of several
reasons:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Both have economic aspects that can create clear advantages for one
  team in a round, given how prior rounds played out. CoD has no such
  equivalent, except for perhaps ``score streaks'', which infrequently
  occur.
\item
  Teams play either offense or defense for many consecutive rounds. (The
  former has a 15-15-1 format and the latter uses a 12-12-1 format.) On
  the other hand, teams in CoD SnD rotate roles every round, analogous
  to a 1-1-1-1-1-1-1 format for home advantage in best-of-seven series
  for professional sports like the MLB, NBA, and NHL.\footnote{1-1-1-1-1-1-1
    is not used today in these leagues, but it was at least once in
    each.} While theoretically one might be able to account for
  inter-round win percentage relationships in any kind of format, such
  as a 5-5-1, the rotation of team sides every round is convenient for
  convincing ourselves that rounds could reasonably be modeled as i.i.d.
  Bernoulli trials.
\item
  Both Counter-Strike and Valorant have overtime rules---a team must win
  by two rounds---which can make end-of-series sequences difficult to
  model properly. CoD SnD does not have overtime rules.
\end{enumerate}

To our knowledge, there is no existing public statistical research on
the CoD SnD format, beyond descriptive analysis on social media. While
intra-round trends may be more directly applicable to teams looking for
an advantage on their competition, broader investigation of a concept
like the hot-hand fallacy in a sport where it has not yet been
invetigated---particularly one that is less subjected to factors that
may be difficult to control for, e.g.~weather---should be useful as an
inspiration for future researchers.

@@

In a best-of-\(s\) format, assuming a constant round win probability
\(\phi\) (where \(0 \leq \phi \leq 1\))\footnote{If \(\phi > 0.5\), we
  can say that this is the stronger team.} for one team\footnote{\(\phi\)
  and other symbols are selected so as to reserve symbols like \(p\) for
  usage in other contexts without causing confusion for the reader.
  Upper-case symbol, e.g.~\(\Phi\), are consistently used in this paper
  to represent proportions, while lower-case symbols corresponding to
  upper-case symbols, e.g.~\(\phi\), are used to represent Bernoulli
  trial success rates. ``Hats'', e.g.~\(\hat{\Phi}\), are used to convey
  expectations, while bare symbols symbols convey observational data.},
the expected proportion of series ending in \(r\) rounds (\(r \leq s\))
is given by Equation \ref{eq:series-length}.

@@

in which \(\hat{\Phi}(r)\) and \(\Phi(r)\) represent the expected and
observed round win proportions respectively, and
\(R = [6, 7, 8, 9, 10, 11]\) for CoD SnD. The null hypothesis is that
the test statistic is chi-squared distributed, where the number of
degrees of freedom is \(\vert R \vert -1 = 5\) for Cod SnD.

@@

Let us now consider round win proportions immediately following a streak
of \(k\) wins, given that the series lasts \(r\) rounds. Removing our
knowledge of a streak, we might model the Bernoulli round win
probability as

\begin{equation}\label{eq:pwr}
p_0(\text{win} | r) = p^{+|r}_0 := \begin{cases} 
  \frac{m}{r} & \text{team wins series}, \\ 
  \frac{r - m}{r} & \text{team loses series},
\end{cases}
\end{equation}

@@

As implied by MS's Theorem 1, we should expect the proportion of round
wins immediately following a streak of \(k\) rounds wins for a series
lasting \(r\) rounds, \(\hat{P}^{+|k,r}_{MS}\), to be strictly less than
\(p^{+|r}\). However, we should not mirror the equal proportions
hypothesis testing performed by MS.

Our context is fundamentally different from that of MS, who focus on
longitudinal data in controlled settings, unaffected by
opposition.\footnote{GVT also perform statistical tests on shots from
  players in live games, i.e.~``observational'' data, but they note that
  their findings are likely affected by player shot selection in the
  face of tactics by the opposing team.} The number of trials is fixed
in their experimental designs, but in CoD SnD, the number of rounds
played is determined as a function of the max number of possible rounds
(\(s\)) and whether or not the team wins the series. The Bernoulli trial
success probability, i.e.~single round win percentage in CoD SnD, is not
independent of the opponent. Consequently, a statistical test of the
difference in \(\hat{P}^{+|k,r}\) and \(\hat{P}^{-|kr}\), as performed
by MS to evaluate their hypothesis regarding post-streak success rate,
is not completely appropriate, although useful as a reference.

@@

One can consider another form of the expected proportion of rounds won
immediately after a streak of \(k\) round wins in a best-of-\(s\) series
given the final length of the series (\(r\) rounds), the ``notional''
proportion \(\hat{P}^{+|k,r}_0\). The Bernoulli round win probability
underlying \(\hat{P}^{+|k,r}_0\) is

\begin{equation}\label{eq:pwkr}
p_0(\text{win} | k, r) = p^{+|k,r}_0 := \begin{cases}
  \frac{m - k}{r - k} & \text{team wins series}, r > m; \\
  1 & \text{team wins series}, r = m; \\
  \frac{s - m - k}{r - k} & \text{team loses series}, r > m + k; \\
  0 & \text{team loses series}, r = m + k,
\end{cases}
\end{equation}

where \(r \in R\) for \(m\) from Equation \ref{eq:m}.

We can perform a two-tailed binomial test to evaluate the null
hypothesis \(\omega = \omega_0\) for the observed probability of success
\(\omega\) and a user-specified \(\omega_0\) (where
\(0 \leq \omega_0 \leq 1\)). If there are \(r^+\) observed successes in
a sample of \(r\) trials, the probability of arriving at this expected
number of successes is

\begin{equation}\label{eq:binom}
\binom {r}{r^+} \omega^{r^+}(1-\omega)^{r-r^+}.
\end{equation}

We should expect that there should be \(r * \omega_0\) successes.

Treating the notional proportion \(\hat{P}^{+|k,r}_0\) as the null
\(\omega_0\) and plugging in the observed proportion \(P^{+|k,r}\) for
\(\omega\) in Equation \ref{eq:binom} (treating proportions as
probabilities), we can evaluate the null hypothesis that the observed
proportion of post-streak round wins is equal to the notional
proportion, given streak length \(k\) and series length \(r\). If we can
reject this null hypothesis, then we can consider team momentum,
represented by post-streak success, plausible.

As a reference, we can perform the same binomial test for the MS's
streak-selection-adjusted proportion, \(\hat{P}^{+|k,r}_{MS}\). However,
given the caveats mentioned before regarding applying MS's theorem to
our setting, the results of such binomial tests should be heeded with
caution.

@@

\begin{equation}\label{eq:pwkri}
p_0(\text{win} | k, r, i) = p^{+|k,r,i}_0 = \begin{cases}
  \frac{m - k}{r - k} & \text{team wins series}, i \neq r; \\
  1 & \text{team wins series}, i = r; \\
  \frac{s - m - k}{r - k} & \text{team loses series}, i \neq r, r > m + k; \\
  0 & \text{team loses series}, i = r \text{ or } r = m + k.
\end{cases}
\end{equation}

Again, we can apply a binomial test to evaluate the hypothesis that the
expected proportion--\(\hat{P}^{+|k,r,i}_0\) and
\(\hat{P}^{+|k,r,i}_{MS}\) separately--is equal to the observed
proportion \(P^{+|k,r,i}\).

@@

Stepping back from the perspective of a single team's round win
probability when streaking, one can attempt to detect the hot hand
phenomenon more broadly with a Wald-Wolfowitz runs test. Under the null
hypothesis, the number of runs in a sequence of \(r\) trials is a random
variable that can take on values \(+\) or \(-\) and arrive at \(r^+\)
successes (\(r^- = r - r^+\) failures), with the following mean \(\mu\)
and variance \(\sigma^2\):

\begin{equation}\label{eq:ww}
\mu = \frac{2r^{+}r^{-}}{r} + 1, \sigma^2 = \frac{(\mu-1)(\mu-2)}{r-1}.
\end{equation}

To incorporate our findings regarding constant round win probability, We
can specify that \(\Pr(r^+) = \phi\) (and, conversely, that
\(\Pr(r^-) = 1 - \phi\)).

One can subset the observed series sequences, \(\zeta(r)\), to those
that violate the null hypothesis for the runs test and perform a test of
equal proportions, where the null is that the observed proportion of a
sequence relative to all possible sequences, \(P(\zeta) = P^\zeta\), is
equal to the expected proportion of the sequence relative to all
possible sequences, \(\hat{P}^\zeta\). The test statistic is

\begin{equation}\label{eq:prop}
z = \frac{P^\zeta - \hat{P}^\zeta}{\sqrt{P^\zeta_{\delta} (1 - P^\zeta_{\delta}) (1 / N^\zeta + 1 / \hat{N}^\zeta)  } }
\end{equation}

where

\[
P^\zeta_{\delta} = \frac{P^\zeta - \hat{P}^\zeta}{N^\zeta - \hat{N}^\zeta}
\]

and where \(N^\zeta\) and \(\hat{N}^\zeta\) are the observed and
expected number of sequences respectively. The null hypothesis is that
\(z\) is normally distributed. If \(z > +z_{(1-\alpha)/2}\) or
\(z < -z_{(1-\alpha)/2}\) at a significance level \(\alpha\), we can
reject the null hypothesis for such sequences and build an argument in
support of the hot hand effect.

@@

Using Equation \ref{eq:chi-squ}, we find that \(\chi^2 = 16.0\)
(\(p\)-value of 0.0068) for \(\phi_0 = 0.5\). Thus, we can comfortably
reject the constant probability hypothesis for the null
\(\phi_0 = 0.5\), even at a confidence level of \(\alpha = 0.01\).

Table \ref{tbl:prob-series-lasts-r-rounds} shows the expected series
lasting \(r\) rounds, the expected proportion of series given
\(\phi_0 = 0.5\), \(\hat{\Phi}_0(r)\), and the observed proportions,
\(\Phi(r)\).

\begin{table}
\caption{The expected proportion of CoD SnD series lasting $r$ rounds, $\hat{\Phi}_0(r)$, under the assumption that each team has a constant round win probability $\phi_0 = 0.5$. Additionally, the observed frequencies for CoD SnD shown as a count $N(r)$ and as a proportion $\Phi(r)$ of all series ($\sum_{r \in R} N(r)$, where $r \in R = [6, 7, 8, 9, 10, 11]$).}

@@

Table \ref{tbl:mosteller-methods-results} shows the alternate values for
the constant round win probability that we find when applying the three
methods suggested by Mosteller (1952). Each is approximately equal to
0.575. When applying Equation \ref{eq:chi-squ}, each results in a
\(\chi^2\) value for which we cannot reject the constant probability
null hypothesis. Note that rejection of the null hypothesis is the
baseline and failure to reject the null is the non-trivial result,
contrary to most statistical analysis.

@@

Observing that the null hypothesis acceptance shown in Table
\ref{tbl:mosteller-methods-results}, along with the observation that
\(\hat{\Phi}_2(r)\) reasonably matches \(\Phi(r)\) (in comparison to
\(\hat{\Phi}_0(r)\)), we can say that the constant round win probability
assumption is valid in CoD SnD series with the appropriate choice of
\(\phi\) (\(\approx 0.575\)).

@@

The relative frequency of the expected proportions are based on 10,000
simulations using the constant round probability \(\phi_2 = 0.575\). A
test for the difference between the observed and expected proportions of
all sequences, \(P^{\zeta r}\) and \(\hat{P}^{\zeta r}\) respectively,
indicates that the null hypothesis---that the two proportions are
equal---cannot be rejected for any of the 25 significant sequences.

>>>>

In professional CoD SnD, the two teams\footnote{In the 2020 season,
  teams played with five players each; in the 2021 and 2022 seasons,
  teams played with four players each.} take turns playing offense and
defense every round. The first to six round wins (best-of-11 round
format) is declared the series winner.\footnote{A short list of rules
  that govern timing follows.} A round can end in one of four ways:

\begin{itemize}
\tightlist
\item
  Each round has a 90 second time limit, not counting the potential
  extension granted if a bomb is planted
\item
  A bomb plant takes five seconds. The timer resets if the player stops
  planting site prior to completing it.
\item
  A bomb defuse takes seven seconds. The timer resets if the player
  ``drops'' the bomb.
\item
  Once the bomb is planted, the round timer is set to 45 seconds.
\end{itemize}

@@

The data set consists of all SnD matches played in major tournaments and
qualifiers during the past three years\footnote{Since the release of
  ``Call of Duty: Modern Warfare'' in fall 2019, professional CoD has
  been orchestrated by the 12-franchise CoD League (CDL). The CDL has
  completed three year-long ``seasons'' as of August 2022.}, totaling
7,792 rounds across 852 series.\footnote{288 of the series occur in
  major tournaments, which are considered to be more competitive than
  the qualifiers since they are played in person (COVID-permitting),
  eliminating randomness due to online latency and ping. The qualifiers
  are played online.} Data has been collected in spreadsheets by
community member ``IOUTurtle''.\footnote{Raw data:
  \url{https://linktr.ee/CDLArchive}. Cleaned data:
  \url{https://github.com/\%7Bauthor\%7D/\%7Brepo\%7D/blob/master/data/\%7Bfile\%7D.csv}.}

CoD is fairly unique compared to other esports in that it runs on an
annual lifecycle, with releases coming in the late fall. A new game
under the same brand---``Call of Duty''---is published every year by a
rotating set of developers. Each new game bears resemblance to past
ones, often introducing relatively small variations (``improvements'')
to graphics, game modes, and other facets of gameplay. During the CDL
era, the games released have been ``Modern Warfare'' (2020 season),
``Cold War'' (2021 season) and ``Vanguard'' (2022 season).

Sweeps (6-0 series) make up 4.7\% of all series, while 6-5 series make
up 23.7\% of series. The observed offensive round win
percentage\footnote{We use the terms ``percentage'', ``probability'',
  and ``proportion'' synonymously throughout this paper when discussing
  round win rates. For the most part, we restrict ``percentage'' usage
  to observed values and ``probability'' usage to expected values, and
  we use ``proportion'' to refer to either observed or expected values.}
across all rounds is \(\bar{\tau}\) = 47.8\%.\footnote{Offensive round
  win percentage has been nearly constant across the three games during
  the CDL era: 47.2\% in Modern Warfare; 47.9\% in Cold War; and 48.1\%
  in Vanguard} Table \ref{tbl:o-win-prop-by-series-state} shows
offensive round win percentages by series ``state'', i.e.~the number of
round wins by each team prior to an upcoming round. Offensive round win
percentage is not quite constant, although never veers more than 10\%
from \(\bar{\tau}\).

@@

The professional competition format balances the frequency with which
teams play offense or defense to start a series, to the extent possible.
Thus, win percentages in certain states are subjected to minimal
selection bias, i.e.~bias due to better teams playing defense more
frequently.\footnote{At tournaments, higher-seeded (``better'') teams
  choose whether they want to start on offense in the SnD series, played
  as the second or fifth game mode of the best-of-five matchup. Their
  opponent is assigned to start on offense in the SnD slot not chosen.
  Anecdotally, the better team tends to choose to play defense in the
  first round of the SnD series played as the second game mode in the
  best-of-five matchup, although this choice is not consistent across
  teams, or even with the same team over time.}

@@

This format has minimized the range of rates at which teams start series
playing offense. The highest rate at which a team has started their
series playing defense across all tournament and qualifier SnD series in
a single season is 66\% (31 of 47). The lowest is 38\% (16 of 42).

@@

There have been a handful of studies of the distribution of games played
in a series of a professional sport. Most assume a constant probability
\(\phi\) of a given team winning a game in the series, regardless of the
series state. Mosteller (\protect\hyperlink{ref-mosteller1952}{1952})
observed that the American League had dominated the National League in
MLB World Series matchups through 1952, implying that games should not
be modeled with a constant \(\phi_0 = 0.5\). Mosteller proposed three
approaches for identifying the optimal constant probability value of the
stronger team in the World Series, finding \(\hat{\phi} \approx 0.65\)
in each case: (1) solve for \(\phi\) from the observed average number of
games won by the loser of the series, which he called the ``method of
moments'' approach; (2) maximize the likelihood that the sample would
have been drawn from a population in which the probability of a team
winning a game is constant across the series (i.e.~maximum likelihood),
and (3) minimize the chi-squared goodness of fit statistic \(\chi^2\) as
a function of \(\phi\).

Chance (\protect\hyperlink{ref-chance2020}{2020}) re-examined the
constant probability notion in the MLB World Series (1923--2018), the
NBA Finals (1951--2018), and the NHL Stanley Cup (1939--2018). Chance
found strong evidence against the null hypothesis of \(\phi_0 = 0.5\) in
the MLB and NHL championship series when applying Mosteller's first and
second methods.

The concept of momentum goes hand-in-hand with a discussion of the
nature of series outcomes.\footnote{We often use ``streaks'' and
  momentum interchangeably, but as Steeger et al.
  (\protect\hyperlink{ref-steeger2021}{2021}) note, momentum implies
  dependence between events, whereas streaking does not.} Two opposing
fallacies are observed in the context of momentum: the ``gambler's
fallacy'' (negative recency) and the ``hot hand fallacy'' (positive
recency). Per Ayton et al. (\protect\hyperlink{ref-ayton2004}{2004}),
negative recency is ``the belief that, for random events, runs of a
particular outcome \ldots{} will be balanced by a tendency for the
opposite outcome'', while positive recency is the expectation of
observing future results that match recent results.

Studying both player streaks and team streaks in basketball, in both
observational and controlled settings, Gilovich et al.
(\protect\hyperlink{ref-gilovich1985}{1985}), henceforth GVT, did not
find evidence for the hot hand phenomenon. However, Miller and Sanjurjo
(\protect\hyperlink{ref-miller2018}{2018}), henceforth MS, provided a
framework for quantifying streak selection bias, which effectively works
in the manner posited by the gambler's fallacy. Specifically, MS stated
that a ``bias exists in a common measure of the conditional dependence
of present outcomes on streaks of past outcomes in sequential data''
implying that, under i.i.d. conditions, ``the proportion of successes
among the outcomes that immediately follow a streak of consecutive
successes is expected to be strictly less than the underlying
(conditional) probability of success''. When applying streak selection
bias to GVT's data, MS came to the opposite conclusions as GVT.

@@

Despite the plethora of existing research on games played in a series
and momentum in sports, these topics have yet to be investigated
heavily, if at all, in esports. Work has been done to examine
intra-round, second-by-second win probability in other FPS titles such
as Counter-Strike (\protect\hyperlink{ref-xenopoulos2022}{Xenopoulos,
Freeman, and Silva 2022}) and Valorant
(\protect\hyperlink{ref-derover2021}{DeRover 2021}), both of which are
round-based like CoD SnD. However, considering Counter-Strike and
Valorant specifically, research on round-level trends seems
non-existent, perhaps for one of several reasons:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Both have economic aspects that can create clear advantages for one
  team in a round, given how prior rounds played out. CoD has no such
  equivalent, except for perhaps ``score streaks'', which infrequently
  occur.
\item
  Teams play either offense or defense for many consecutive rounds.
  (Counter-Strike has a 15-15-1 format and Valorant has a 12-12-1
  format.) On the other hand, teams in CoD SnD rotate attacking and
  defending roles every round, minimizing the effect of map asymmetry on
  streaks of wins.\footnote{Professional CoD has a set of maps (at least
    three, depending on the season) on which teams choose to play SnD.
    Most are asymmetric, introducing slight advantages for the offense
    or defense that do not exist on other maps. Counter-Strike and
    Valorant also have their own sets of maps for their analogous SnD
    modes.}
\item
  Both Counter-Strike and Valorant have overtime rules---a team must win
  by two rounds---which can make end-of-series sequences difficult to
  model properly. CoD SnD does not have overtime rules.
\end{enumerate}

To our knowledge, there is no existing public statistical research on
the CoD SnD format, beyond descriptive analysis on social media. While
intra-round trends may be more directly applicable to teams looking for
an advantage on their competition, broader investigation of a concept
like the hot-hand fallacy in a sport where it has not yet been
investigated---particularly one that is less subjected to factors that
may be difficult to control for, e.g.~weather---should be useful as an
inspiration for future researchers.

@@

In a best-of-\(s\) format, assuming a constant round win probability
\(\phi\) (where \(0 \leq \phi \leq 1\)) for a given team, the expected
proportion of series ending in \(r\) rounds (where \(r \leq s\)) is
given by Equation \ref{eq:series-length}.

@@

where \(\hat{\Phi}(r)\) and \(\Phi(r)\) represent the expected and
observed round win proportions respectively, and where
\(R = [6, 7, 8, 9, 10, 11]\) for CoD SnD. The null hypothesis is that
the test statistic is chi-squared distributed, where the number of
degrees of freedom is \(\vert R \vert -1 = 5\) for Cod SnD. We use a
significance level of \(\alpha = 0.05\) for this test and all others in
this paper.

@@

Let us now consider round win proportions immediately following a streak
of \(k\) wins, given that the series lasts \(r\) rounds. Removing our
knowledge of a streak, we might expect the win proportion for a round to
be

\begin{equation}\label{eq:pwr}
\hat{P}_0(\text{win} | r) = \hat{P}^{+|r}_0 := \begin{cases} 
\frac{m}{r} & \text{team wins series}, \\ 
\frac{r - m}{r} & \text{team loses series},
\end{cases}
\end{equation}

@@

As implied by MS's Theorem 1, we should expect the proportion of round
wins immediately following a streak of \(k\) rounds wins for a series
ending in \(r\) rounds, \(\hat{P}^{+|k,r}_{MS}\), to be strictly less
than the observed proportion of round wins given the series length, but
not given knowledge of the potential existence of a streak, \(P^{+|r}\),
i.e.~the analogue of the expected proportion in Equation \ref{eq:pwr}
for the observed data. However, our context is fundamentally different
from that of MS, who evaluate data in controlled settings, unaffected by
opposition.\footnote{GVT also perform statistical tests on shots from
  players in live games, i.e.~``observational'' data, but they note that
  their findings are likely affected by player shot selection influenced
  by the opposing team's defensive strategy. This is similar to our
  setting.}

The number of trials is fixed in their experimental designs; in CoD SnD,
however, the number of rounds played is determined as a function of the
max number of possible rounds (\(s\)) and whether or not the team wins
the series. The single round win percentage in CoD SnD---analogous to a
player's shooting percentage in MS's analysis of basketball players---is
not independent of the opponent. Consequently, the i.i.d. assumption of
MS's theorem regarding post-streak success may or may not be valid.
Nonetheless, we still leverage \(\hat{P}^{+|k,r}_{MS}\) as a reference
for our results.

@@

Let us consider another form of the expected proportion of rounds won
immediately after a streak of \(k\) round wins in a best-of-\(s\) series
ending in \(r\) rounds: the ``notional'' proportion

\begin{equation}\label{eq:pwkr}
\hat{P}_0(\text{win} | k, r) = \hat{P}^{+|k,r}_0 := \begin{cases}
\frac{m - k}{r - k} & \text{team wins series}, r > m; \\
1 & \text{team wins series}, r = m; \\
\frac{s - m - k}{r - k} & \text{team loses series}, r > m + k; \\
0 & \text{team loses series}, r = m + k,
\end{cases}
\end{equation}

where \(r \in R\) and \(m\) from Equation \ref{eq:m}.

We can perform a two-tailed binomial test to evaluate the null
hypothesis \(\omega = \omega_0\) for the observed proportion of
successes, \(\omega\), and a null proportion, \(\omega_0\) (where
\(0 \leq \omega_0 \leq 1\)). The probability of exactly \(r^+\)
successes (and \(r^- = r - r^+\) failures) in a sample of \(r\) trials
is

\begin{equation}\label{eq:binom}
\binom {r}{r^+} \omega^{r^+}(1-\omega)^{r^-},
\end{equation}

where \(\omega = \frac{r^+}{r}\) is the observed proportion of
successes. If we can reject the null hypothesis---treating \(P^{+|k,r}\)
and \(\hat{P}^{+|k,r}_0\) as \(\omega\) and \(\omega_0\)
respectively---then we might consider team momentum plausible. Since we
perform this test for many combinations of \(k\) and \(r\), we adjust
the \(p\)-values with the Benjamini and Yekutieli (BY) false discovery
procedure (\protect\hyperlink{ref-benjamini2001}{2001}).

As a reference, we can perform the same binomial test for MS's
proportion, \(\hat{P}^{+|k,r}_{MS}\), which adjusts for streak selection
bias. However, given the caveats mentioned before regarding the
application of MS's theorem to our setting, the results of such binomial
tests should be heeded with caution.

@@

\begin{equation}\label{eq:pwkri}
\hat{P}_0(\text{win} | k, r, i) = \hat{P}^{+|k,r,i}_0 := \begin{cases}
\frac{m - k}{r - k} & \text{team wins series}, i \neq r; \\
1 & \text{team wins series}, i = r; \\
\frac{s - m - k}{r - k} & \text{team loses series}, i \neq r, r > m + k; \\
0 & \text{team loses series}, i = r.
\end{cases}
\end{equation}

Again, we can apply a binomial test to evaluate the hypothesis that the
expected proportion, \(\hat{P}^{+|k,r,i}_0\), is equal to the observed
proportion, \(P^{+|k,r,i}\).

@@

We can attempt to detect the hot hand phenomenon more broadly by
focusing on non-random sequences identified by the Wald-Wolfowitz runs
test. Note that the probability of success and failure need not be
equal; thus, we can incorporate our findings regarding constant round
win probability, specifying that the probability of success is
\(\phi_2 = 0.575\).

Under the null hypothesis for the Wald-Wolfowitz runs test, the number
of runs in a sequence of \(r\) trials is a random variable that can take
on values \(+\) or \(-\) and arrive at \(r^+\) successes (and \(r^-\)
failures), with the following mean \(\mu_r\) \footnote{We calculate
  \(\mu\) and \(\sigma^2\) for each number of rounds \(r \in R\).} and
variance \(\sigma_r^2\):

\begin{equation}\label{eq:ww}
\mu_r = \frac{2r^{+}r^{-}}{r} + 1, \sigma_r^2 = \frac{(\mu_r-1)(\mu_r-2)}{r-1}.
\end{equation}

The null hypothesis is that the test statistic,

\begin{equation}\label{eq:wwz}
z_r = \frac{r^* - \mu_r}{\sigma_r},
\end{equation}

where \(r^*\) is the number of runs, is normally distributed.

We can subset the observed series sequences, \(\zeta_j\), to those that
violate the null hypothesis for the runs test and perform a test of
equal proportions, where the null hypothesis is that the observed
proportion of a sequence relative to all possible sequences,
\(P^{\zeta}_j\) ( \(\frac{N^\zeta_j}{\sum_j N^\zeta }\)), is equal to
the expected proportion of the sequence relative to all possible
sequences, \(\hat{P}^\zeta_j\). The test statistic is

\begin{equation}\label{eq:prop}
z^\zeta_j = \frac{P^\zeta_j - \hat{P}^\zeta_j}{\sqrt{P^\zeta_{j,\delta} (1 - P^\zeta_{j,\delta}) (1 / N^\zeta_j + 1 / \hat{N}^\zeta_j)  } }
\end{equation}

where \(N^\zeta_j\) and \(\hat{N}^\zeta_j\) respectively represent the
observed and expected number of occurrences of sequence \(\zeta_j\), and
where
\(P^\zeta_{j,\delta} = (P^\zeta_j - \hat{P}^\zeta_j) / ({N^\zeta_j - \hat{N}^\zeta_j})\).
The null hypothesis is that the test statistic is normally distributed.
If we can reject the null hypothesis for such sequences, we can build an
argument in support of the hot hand effect.\footnote{As with the
  binomial tests for the notional and MS post-streak proportions, we
  adjust the \(p\)-values with the BY correction.}

@@

Using Equation \ref{eq:chi-squ}, we find that \(\chi^2 = 16.0\)
(\(p\)-value of \textless0.01) for \(\phi_0 = 0.5\). Thus, we can
comfortably reject the constant probability hypothesis for the naive
\(\phi_0 = 0.5\).

Table \ref{tbl:prob-series-lasts-r-rounds} shows the expected proportion
of series ending in \(r\) rounds given \(\phi_0 = 0.5\),
\(\hat{\Phi}_0(r)\), as well as the observed proportions, \(\Phi(r)\).

\begin{table}
\caption{The observed frequencies of CoD SnD series ending in $r$ rounds shown as a count $N(r)$ and as a proportion $\Phi(r)$ of all series ($\sum_{r \in R} N(r)$, where $r \in R = [6, 7, 8, 9, 10, 11]$). Additionally, the expected proportion, $\hat{\Phi}_0(r)$, under the assumption that each team has a constant round win probability $\phi_0 = 0.5$.}

@@

Table \ref{tbl:mosteller-methods-results} shows the alternate values for
the constant round win probability assumption, each approximately equal
to 0.575. When applying Equation \ref{eq:chi-squ}, each results in a
\(\chi^2\) value for which we cannot reject the constant probability
null hypothesis.

@@

The observation that \(\hat{\Phi}_2(r)\) reasonably matches \(\Phi(r)\)
supports the implication from the failure to reject the null hypothesis
for \(\phi_2\) (shown in Table \ref{tbl:mosteller-methods-results}): the
constant round win probability assumption can be valid for CoD SnD with
the appropriate choice of \(\phi\) (\(\approx 0.575\)).

@@

The expected proportion of occurrences of each sequence \(\zeta_j\) is
based on 10,000 simulations using the constant round probability
\(\phi_2 = 0.575\). A test for the difference between the observed and
expected proportions of all sequences, \(P^{\zeta}_j\) and
\(\hat{P}^{\zeta}_j\) respectively, indicates that the null
hypothesis---that the two proportions are equal---cannot be rejected for
any of the non-random sequences.
